---
title: "Handling author and affiliation with a RMarkdown file."
output: 
  word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("tidyverse")
library("flextable")

```

## Assuming we have a dataframe with the following author information (assuming the author order is the same as it should be on the paper):

- name
- affiliation 1, 2, 3, ...
- address of the affiliation 1, 2, 3, ...

## Each affiliation and address should be in separate columns.

```{r, warning=FALSE, message=FALSE, echo=TRUE}

df <- data.frame(
  Name = c("Georg J. A. Hähn", "Max Mustermann"),
  `Affiliation1` = c("MLU", "iDiv"),
  `Address1` = c("Halle (Saale), Germany", "Leipzig, Germany"),
  `Affiliation2` = c("iDiv", "Musteruni"),
  `Address2` = c("Leipzig, Germany", "Musterstadt, Musterland"),
  `Affiliation3` = c("Uni Bologna", NA),
  `Address3` = c("Bologna, Italy", NA)
)


flextable(df[,1:4]) %>% 
  theme_zebra() %>% 
  autofit()

```

## Now we combine the affiliation with the address:
```{r, warning=FALSE, message=FALSE, echo=TRUE}
df1 <- df %>% 
  mutate(
  Affiliation1 = ifelse(!is.na(`Affiliation1`), paste0(`Affiliation1`, ", ", `Address1`), NA),
  Affiliation2 = ifelse(!is.na(`Affiliation2`), paste0(`Affiliation2`, ", ", `Address2`), NA),
  Affiliation3 = ifelse(!is.na(`Affiliation3`), paste0(`Affiliation3`, ", ", `Address3`), NA)
  ) %>% 
  dplyr::select(Affiliation1, Affiliation2, Affiliation3, Name)

flextable(df1[,1:4]) %>% 
  theme_zebra() %>% 
  autofit()
```

## Here, we bring the affiliation to one column and keep the original order of the authors.
```{r, warning=FALSE, message=FALSE, echo=TRUE}
# Get affiliations while keeping the original author order
affil <- gather(df1, key = "Level", value = "Affiliation", -c(Name)) %>% 
  filter(stringr::str_detect(Level, "Affil")) %>% 
  arrange(match(Name, df1$Name))

flextable(affil) %>% 
  theme_zebra() %>% 
  autofit()
```

## The affiliations are ordered, therefore we can simply add the numbers 1 to n(affiliations)
```{r, warning=FALSE, message=FALSE, echo=TRUE}
affil_order <- data.frame(Affiliation = na.omit(unique(affil$Affiliation)),
                          Number = 1:length(na.omit(unique(affil$Affiliation))))

flextable(affil_order) %>% 
  theme_zebra() %>% 
  autofit()
```

## Link the number to the affiliation in the author list, and combine all numbers for each author in one column:
```{r, warning=FALSE, message=FALSE, echo=TRUE}
authors_numbers = affil %>% 
  right_join(affil_order %>% na.omit(), by = "Affiliation") %>% 
  group_by(Name) %>% 
  summarize(Affil = paste(Number, collapse = ",")) %>% 
  arrange(match(Name, df1$Name))

flextable(authors_numbers) %>% 
  theme_zebra() %>% 
  autofit()
```

## For the authors we can simply create a long string with all names and the numbers as superscript. For the first author we add the * as we want to indicate that he is the first author of the paper.
```{r, warning=FALSE, message=FALSE, echo=TRUE}
authorlist <- ""

for (i in 1:length(authors_numbers$Name)) {

  str <- paste0(authors_numbers[i, "Name"], "^", ifelse(i == 1, paste0(authors_numbers[i, "Affil"], ",\\*"), 
                                                        authors_numbers[i, "Affil"]), 
                       "^", ifelse(i == length(authors_numbers$Name), "",  ", "), sep = "")
  authorlist <- paste0(authorlist, str, sep = "")
}

```
`r authorlist`

## For the affiliation list we do the same as for the authors but we sperate them with a linebreak instead of a comma.

```{r, warning=FALSE, message=FALSE, echo=TRUE}
affillist <- "\\* Corresponding author   \n   \n"

for (i in 1:length(affil_order$Affiliation)) {
  a <- paste0("^", affil_order[i, "Number"], "^", " ", affil_order[i, "Affiliation"], sep = "\n")
  affillist <- paste0(affillist, a)
}

```

## To print the affiliation list we need to slightly modify the `cat` function:
```{r, results='asis', echo=TRUE}
mycat <- function(text){
  cat(gsub(pattern = "\n", replacement = "  \n", x = text))
}
mycat(affillist)
```


## Done. The final result:

`r authorlist`

```{r, results='asis', echo=FALSE}
mycat <- function(text){
  cat(gsub(pattern = "\n", replacement = "  \n", x = text))
}
mycat(affillist)
```

